---
title: "Shockbox Analysis"
author: "Isla Weber"
date: "5/6/2021"
output: html_document
---
load necessary libraries
```{r}
library(ggplot2)
library(readr)
library(tidyr)
```


# Import data

To import, *replace the filepath* below with the appropriate path location. (Run getwd() in the console to see where the filepath should start from, should be folder where this file is saved). Must be a csv file. 

To analyze data from all cohorts, go to https://docs.google.com/spreadsheets/d/1oTD1mxHx49RlTYxX7h7U56rEziUqdsdb/edit#gid=1158166666 and create a csv of the first tab in the sheet. 

```{r}
data_raw <- read_csv("all_cohorts_freezing_temp.csv")
head(data_raw)

# for now I just want to look at recall 1-5 because they are completed, so I remove the others:
data <- data_raw[data_raw$day != "recall6" & data_raw$day != "recall7" & data_raw$day != "NA",]
```

# Create Plots
```{r}
# This line is used to tell R what order we want the days to go in on the graph:
data$day <- factor(data$day, levels = c("sefla", "seflb", "recall1", "recall2", "recall3", "recall4", "recall5"))

levels(data$day) <- c("SEFL A", "SEFL B", "retrieval 1", "2", "3", "4", "5")

# Create a list of all mouse IDs in the data:
allMice <- c(unique(data$id))

# Create a list of the control mice IDs:
controlMice <- c(unique(data$id[data$condition == "control"]))

'%!in%' <- function(x,y)!('%in%'(x,y))

seflMice <- allMice[allMice %!in% controlMice]

data$condition[data$id %in% controlMice] <- "control"
data$condition[data$id %in% seflMice] <- "ptsd"
```

The ggplot function loads the data and plots. It is very flexible, you can add a lot of features. To find a way to change or add to this, search something like "ggplot R barplot change colors" and you will find examples of code to try and incorporate. 

The aes argument allows you to indicate the groups (fill), y axis (freezing), and x axis(day)

geom_bar adds the bar plot, geom_line adds the lines, geom_point adds the dots.

```{r}
#select the colors which will be associated with each group using universal color codes
group.colors = c(control = "azure4", ptsd = "darkorange2")

#This will plot the grouped bar plot
groupedbar <- ggplot(data, aes(fill=condition, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.9) +
    ggtitle("SEFL Freezing - All mice (N=54)") + ylab("% freezing") + ylim(c(0,100)) +
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors) 


#This will plot the averages in a bar plot then plot the lines for each mouse ID
groupedbar_line <- ggplot(data, aes(fill=condition, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+
    geom_line(aes(x=day, y=freezing, group=id, colour=condition), size=.25) + 
    ggtitle("SEFL Freezing - All mice (N=54)") + ylab("% freezing") + ylim(c(0,100)) +
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)

#This will plot the grouped bar plot plus the individual points
groupedbar_points <- ggplot(data, aes(fill=condition, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.8) +
    geom_point(aes(y=freezing, x=day),
           stat="identity",
           position = position_dodge(width = .9),
           alpha=1,
           size=.75) +
    ggtitle("SEFL Freezing - All mice (N=54)") + ylab("% freezing") +
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)
    
#This will plot just the line plots for each mouse, color coded by group
line_by_id <- ggplot(data=data, aes(x=day, y=freezing, group=id, colour=condition)) + geom_line(size=.25) +
    ggtitle("SEFL Freezing - All mice (N=54)") + ylab("% freezing") +
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)
```
# Plots by sex
```{r}
#MALES
#This will plot the averages in a bar plot then plot the lines for each mouse ID
male_groupedbar_line <- ggplot(data[data$sex=="male",], aes(fill=condition, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+
    geom_line(aes(x=day, y=freezing, group=id, colour=condition), size=.25) + 
    ggtitle("SEFL Freezing - Males (N=42)") + ylab("% freezing") + 
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)


#FEMALES
#This will plot the averages in a bar plot then plot the lines for each mouse ID
female_groupedbar_line <- ggplot(data[data$sex=="female",], aes(fill=condition, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+
    geom_line(aes(x=day, y=freezing, group=id, colour=condition), size=.25) + 
    ggtitle("SEFL Freezing - Females (N=12)") + ylab("% freezing") +
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)
```
# Display Plots
```{r}
groupedbar

groupedbar_line

groupedbar_points

line_by_id

male_groupedbar_line

female_groupedbar_line
```

# separate by resilient vs susceptible and plot

Here we split up the SEFL treated mice based on recall day 5, identifying the top 30 percentile as susceptible and the bottom 30 percentile as resilient.
```{r}
recall5_percentiles_30_90 <- c(quantile(data$freezing[data$day == "recall5"], probs=0.3, na.rm=TRUE), quantile(data$freezing[data$day == "recall5"], probs=0.7, na.rm=TRUE))

resilientMice <- data$id[data$id %in% seflMice & data$day=="recall5" & data$freezing <= recall5_percentiles_30_90[1]]

susceptibleMice <- data$id[data$id %in% seflMice & data$day=="recall5" & data$freezing >= recall5_percentiles_30_90[2]]

data$response[data$id %in% resilientMice] <- "resilient"
data$response[data$id %in% susceptibleMice] <- "susceptible"
data$response[data$id %in% controlMice] <- "control"
```

```{r}

#select the colors which will be associated with each group using universal color codes
response.colors = c(control = "gray75", resilient = "forestgreen", susceptible = "red2")

groupedbar <- ggplot(data[!is.na(data$response),], aes(fill=response, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+ 
    ggtitle("SEFL Freezing - All mice (N=54)") + ylab("% freezing") + 
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=response.colors) + scale_color_manual(values=response.colors)

groupedbar_line <- ggplot(data[!is.na(data$response),], aes(fill=response, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+
    geom_line(aes(x=day, y=freezing, group=id, colour=response), size=.25) + 
    ggtitle("SEFL Freezing - All mice (N=54)") + ylab("% freezing") + 
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=response.colors) + scale_color_manual(values=response.colors)

groupedbar
groupedbar_line
```

Now separate by sex:

```{r}

male_groupedbar <- ggplot(data[!is.na(data$response) & data$sex == "male",], aes(fill=response, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+ 
    ggtitle("SEFL Freezing - Males (N=42)") + ylab("% freezing") +
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=response.colors) + scale_color_manual(values=response.colors)

male_groupedbar_line <- ggplot(data[!is.na(data$response) & data$sex =="male",], aes(fill=response, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+
    geom_line(aes(x=day, y=freezing, group=id, colour=response), size=.25) + 
    ggtitle("SEFL Freezing - Males (N=42)") + ylab("% freezing") + 
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=response.colors) + scale_color_manual(values=response.colors)

male_groupedbar
male_groupedbar_line

female_groupedbar <- ggplot(data[!is.na(data$response) & data$sex == "female",], aes(fill=response, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+ 
    ggtitle("SEFL Freezing - Females (N=12)") + ylab("% freezing") +
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=response.colors) + scale_color_manual(values=response.colors)

female_groupedbar_line <- ggplot(data[!is.na(data$response) & data$sex =="female",], aes(fill=response, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+
    geom_line(aes(x=day, y=freezing, group=id, colour=response), size=.25) + 
    ggtitle("SEFL Freezing - Females (N=12)") + ylab("% freezing") +
    #scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) +
    scale_fill_manual(values=response.colors) + scale_color_manual(values=response.colors)

female_groupedbar
female_groupedbar_line
```

# Scatter plots
```{r}
data_wide <- spread(data[,c('freezing', 'id', 'day')], day, freezing)

reg <- function(x, y, col) abline(lm(y~x), col=col) 

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
 usr <- par("usr"); on.exit(par(usr))
 par(usr = c(0, 1, 0, 1))
 r <- abs(cor(x, y))
 txt <- format(c(r, 0.123456789), digits = digits)[1]
 txt <- paste0(prefix, txt)
 text(0.5, 0.5, txt, cex = 1.1, font = 4)
}

panel.lm =  function (x, y, col = par("col"), bg = NA, pch = par("pch"), 
    cex = 1, col.smooth = "red", span = 2/3, iter = 3, ...)  {
    points(x, y, pch = pch, col = col, bg = bg, cex = cex)
    ok <- is.finite(x) & is.finite(y)
    if (any(ok)) reg(x[ok], y[ok], col.smooth)
}

corrplot_all <- pairs(na.omit(data_wide[,2:8]), panel = panel.lm, main = "Correlation Matrix - Freezing (All mice, N=54)",
    cex = 1.5, pch = 19, col = adjustcolor(4, .4), cex.labels = 2, 
    font.labels = 2, lower.panel = panel.cor)
corrplot_all

corrplot_ptsd <- pairs(na.omit(data_wide[data_wide$id %in% seflMice,2:8]), panel = panel.lm, main = "Correlation Matrix - Freezing (PTSD mice only, N=32)",
    cex = 1.5, pch = 19, col = adjustcolor(4, .4), cex.labels = 2, 
    font.labels = 2, lower.panel = panel.cor)
corrplot_ptsd
```

# Violin Plots!

## by condition

All mice
```{r}
violin_allMice <- data %>%
  ggplot(aes(x=day,
             y=freezing,
             fill=condition))+
  geom_violin(draw_quantiles = .5) + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors) +
    ggtitle("SEFL Freezing - All mice (N=54)") + ylab("% freezing")

violin_allMice
```

Males
```{r}
violin_Males <- data[data$sex =="male",] %>%
  ggplot(aes(x=day,
             y=freezing,
             fill=condition))+
  geom_violin(draw_quantiles = .5) + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors) +
    ggtitle("SEFL Freezing - Males (N=42)") + ylab("% freezing") 

violin_Males
```

Females
```{r}
violin_Females <- data[data$sex =="female",] %>%
  ggplot(aes(x=day,
             y=freezing,
             fill=condition))+
  geom_violin(draw_quantiles = .5) + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors) +
    ggtitle("SEFL Freezing - Females (N=12)") + ylab("% freezing") 

violin_Females
```

## by response

All mice
```{r}
violin_allMice <- data[!is.na(data$response),] %>%
  ggplot(aes(x=day,
             y=freezing,
             fill=response))+
  geom_violin(draw_quantiles = .5) + scale_fill_manual(values=response.colors) + scale_color_manual(values=response.colors) + 
    ggtitle("SEFL Freezing - All mice (N=54)") + ylab("% freezing") 

violin_allMice
```

Males
```{r}
violin_Males <- data[!is.na(data$response) & data$sex =="male",] %>%
  ggplot(aes(x=day,
             y=freezing,
             fill=response))+
  geom_violin(draw_quantiles = .5) + scale_fill_manual(values=response.colors) + scale_color_manual(values=response.colors)+ 
    ggtitle("SEFL Freezing - Males (N=42)") + ylab("% freezing") 

violin_Males
```

Females
```{r}
violin_Females <- data[!is.na(data$response) & data$sex =="female",] %>%
  ggplot(aes(x=day,
             y=freezing,
             fill=response))+
  geom_violin(draw_quantiles = .5) + scale_fill_manual(values=response.colors) + scale_color_manual(values=response.colors) + 
    ggtitle("SEFL Freezing - Females (N=12)") + ylab("% freezing") 

violin_Females
```

