---
title: "Shockbox Analysis"
author: "Isla Weber"
date: "5/6/2021"
output: html_document
---
load necessary libraries
```{r}
library(ggplot2)
library(readr)
library(tidyr)
```


# Import data

To import, *replace the filepath* below with the appropriate path location. (Run getwd() in the console to see where the filepath should start from, should be folder where this file is saved). Must be a csv file formatted with the following column names:

- *freezing* :: percent of time spent freezing, should just be a number

- *id* :: contains the mouse ID (can be letters and/or numbers)

- *day* :: contains the day of the experiment, written as the following: sefla, seflb, recall1, recall2, recall3, recall4, recall5

- *condition* :: contains the mouse condition, written as the following: control, ptsd

*change the cohort variable to current cohort when you analyze!!*
```{r}
data <- read_csv("PTSD1_freezing.csv")
cohort <- "PTSD1"
data_wide <- spread(data, day, freezing)
head(data)
head(data_wide)
```

# Create Plots
This line is used to tell R what order we want the days to go in on the graph
```{r}
data$day <- factor(data$day, levels = c("sefla", "seflb", "recall1", "recall2", "recall3", "recall4", "recall5"
                                        ))
```

The ggplot function loads the data and plots. It is very flexible, you can add a lot of features. To find a way to change or add to this, search something like "ggplot R barplot change colors" and you will find examples of code to try and incorporate. 

The aes argument allows you to indicate the groups (fill), y axis (freezing), and x axis(day)

geom_bar adds the bar plot, geom_line adds the lines, geom_point adds the dots.

```{r}
#select the colors which will be associated with each group using universal color codes
group.colors = c(control = "#263FFF", ptsd = "#D80505")


#This will plot the grouped bar plot
groupedbar <- ggplot(data, aes(fill=condition, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.9) +
    ggtitle("SEFL Freezing") + ylab("% freezing") + ylim(c(0,75)) +
    scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)

#This will plot the averages in a bar plot then plot the lines for each mouse ID
groupedbar_line <- ggplot(data, aes(fill=condition, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.65)+
    geom_line(aes(x=day, y=freezing, group=id, colour=condition)) + 
    geom_point(aes(x=day, y=freezing, group=id, colour=condition))+
    ggtitle("SEFL Freezing") + ylab("% freezing") + ylim(c(0,75)) +
    scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)

#This will plot the grouped bar plot plus the individual points
groupedbar_points <- ggplot(data, aes(fill=condition, y=freezing, x=day)) + 
    geom_bar(position="dodge", stat="summary", fun.y = "mean", alpha=.8) +
    geom_point(aes(y=freezing, x=day),
           stat="identity",
           position = position_dodge(width = .9),
           alpha=1,
           size=.75) +
    ggtitle("SEFL Freezing") + ylab("% freezing") +
    scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)
    
#This will plot just the line plots for each mouse, color coded by group
line_by_id <- ggplot(data=data, aes(x=day, y=freezing, group=id, colour=condition)) + geom_line() + geom_point() +
    ggtitle("SEFL Freezing") + ylab("% freezing") +
    scale_x_discrete(labels=c("sefla" = "Context A", "seflb" = "Context B")) + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)
```

# Display Plots
```{r}
groupedbar

groupedbar_line

groupedbar_points

line_by_id
```
