---
title: "SleepBox Analysis"
author: "Isla Weber"
date: "5/8/2021"
output: html_document
---
```{r}
library(readr)
library(ggplot2)
library(tidyverse)
library(tibble)
```
Load in mouse data
1. Go in and remove the first three rows of the counts file for each mouse (I will code this into the script when I get the chance).
2. Write out the code to read in the csv file for each mouse (filepath starts where this markdown file is located).
```{r}
#cre01a1_baseline <- read_csv("SleepBox/Analysis/5ht2acre01a1baseline_PTSD1_CL_Counts.csv")
#cre01a1_recall5 <- read_csv("SleepBox/Analysis/5ht2acre01a-1recall5_PTSD1_CL_Counts.csv")
cre01a2_baseline <- read_csv("SleepBox/Analysis/5ht2acre01a2baseline_PTSD1_CL_Counts.csv")
cre01a2_recall5 <- read_csv("SleepBox/Analysis/5ht2acre01a-2recall5_PTSD1_CL_Counts.csv")
cre01a3_baseline <- read_csv("SleepBox/Analysis/5ht2acre01a3baseline_PTSD1_CL_Counts.csv")
cre01a3_recall5 <- read_csv("SleepBox/Analysis/5ht2acre01a-3recall5_PTSD1_CL_Counts.csv")
fl05a5_baseline <- read_csv("SleepBox/Analysis/5ht2afl05a5baseline_PTSD1_CL_Counts.csv")
fl05a5_recall5 <- read_csv("SleepBox/Analysis/5ht2afl05a-5recall5_PTSD1_CL_Counts.csv")
fl05a6_baseline <- read_csv("SleepBox/Analysis/5ht2afl05a6baseline_PTSD1_CL_Counts.csv")
fl05a6_recall5 <- read_csv("SleepBox/Analysis/5ht2afl05a-6recall5_PTSD1_CL_Counts.csv")
fl05a7_baseline <- read_csv("SleepBox/Analysis/5ht2afl05a7baseline_PTSD1_CL_Counts.csv")
fl05a7_recall5 <- read_csv("SleepBox/Analysis/5ht2afl05a-7recall5_PTSD1_CL_Counts.csv")
sert26a2_baseline <- read_csv("SleepBox/Analysis/sert26a2baseline_PTSD1_CL_Counts.csv")
sert26a2_recall5 <- read_csv("SleepBox/Analysis/sert26a-2recall5_PTSD1_CL_Counts.csv")
sert26a3_baseline <- read_csv("SleepBox/Analysis/sert26a3baseline_PTSD1_CL_Counts.csv")
sert26a3_recall5 <- read_csv("SleepBox/Analysis/sert26a-3recall5_PTSD1_CL_Counts.csv")

head(cre01a2_baseline)
```

Do a bunch of tedius and not efficient stuff with the data to get the final full dataframe in long format as desired. 
```{r}
library(tidyverse)
    
data <- tibble::lst(cre01a2_baseline, cre01a2_recall5, cre01a3_baseline, cre01a3_recall5, fl05a5_baseline, fl05a5_recall5, fl05a6_baseline, fl05a6_recall5, fl05a7_baseline, fl05a7_recall5, sert26a2_baseline, sert26a2_recall5, sert26a3_baseline, sert26a3_recall5)

ids <- c("cre01a2_baseline", "cre01a2_recall5", "cre01a3_baseline", "cre01a3_recall5", "fl05a5_baseline", "fl05a5_recall5", "fl05a6_baseline", "fl05a6_recall5", "fl05a7_baseline", "fl05a7_recall5", "sert26a2_baseline", "sert26a2_recall5", "sert26a3_baseline", "sert26a3_recall5")

condition <- c("ptsd", "ptsd", "control", "control", "control", "control", "ptsd", "ptsd", "control", "control", "ptsd", "ptsd", "control", "control")

for (id in ids) {
  data[[id]] <- rename(data[[id]], countsperminute = "Cnts/min")
  data[[id]] <- data[[id]] %>% 
  group_by(Hr) %>% 
  summarise(mean = mean(countsperminute, na.rm = TRUE))
} 

data <- tibble::lst(
       "cre01a2_baseline" = mutate(data[["cre01a2_baseline"]], condition = "ptsd", id = "cre01a2", time = "baseline"),
        "cre01a2_recall5" = mutate(data[["cre01a2_recall5"]], condition = "ptsd", id = "cre01a2", time = "recall5"),
        "cre01a3_baseline" = mutate(data[["cre01a3_baseline"]], condition = "control", id = "cre01a3", time = "baseline"),
        "cre01a3_recall5" = mutate(data[["cre01a3_recall5"]], condition = "control", id = "cre01a3", time = "recall5"),
        "fl05a5_baseline" = mutate(data[["fl05a5_baseline"]], condition = "control", id = "fl05a5", time = "baseline"),
        "fl05a5_recall5" = mutate(data[["fl05a5_recall5"]], condition = "control", id = "fl05a5", time = "recall5"),
        "fl05a6_baseline" = mutate(data[["fl05a6_baseline"]], condition = "ptsd", id = "fl05a6", time = "baseline"),
        "fl05a6_recall5" = mutate(data[["fl05a6_recall5"]], condition = "ptsd", id = "fl05a6", time = "recall5"),
        "fl05a7_baseline" = mutate(data[["fl05a7_baseline"]], condition = "control", id = "fl05a7", time = "baseline"),
        "fl05a7_recall5" = mutate(data[["fl05a7_recall5"]], condition = "control", id = "fl05a7", time = "recall5"),
        "sert26a2_baseline" = mutate(data[["sert26a2_baseline"]], condition = "ptsd", id = "sert26a2", time = "baseline"),
        "sert26a2_recall5" = mutate(data[["sert26a2_recall5"]], condition = "ptsd", id = "sert26a2", time = "recall5"),
        "sert26a3_baseline" = mutate(data[["sert26a3_baseline"]], condition = "control", id = "sert26a3", time = "baseline"),
        "sert26a3_recall5" = mutate(data[["sert26a3_recall5"]], condition = "control", id = "sert26a3", time = "recall5"))

full_data <- bind_rows(data, .id = "column_label")
```

Make graphs!
```{r}
#select the colors which will be associated with each group using universal color codes
group.colors = c(control = "#263FFF", ptsd = "#D80505")

pre_line_by_id <- ggplot(full_data[full_data$time == "baseline",], aes(x=Hr, y=mean, group = id, colour =condition)) + geom_line() +
    ggtitle("Sleep-Wake pre trauma") + ylab("Mean Activity") + xlab("Hour of Day (0-23)") + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)+ ylim(c(0,.25))

post_line_by_id <- ggplot(full_data[full_data$time == "recall5",], aes(x=Hr, y=mean, group = id, colour =condition)) + geom_line() +
    ggtitle("Sleep-Wake post trauma") + ylab("Mean Activity") + xlab("Hour of Day (0-23)") + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors)+ ylim(c(0,.25))

pre_line_grouped <- ggplot(full_data[full_data$time == "baseline",], aes(x=Hr, y=mean, group = id, colour =condition)) + stat_summary(aes(group=condition), fun.y=mean, geom="line") + ggtitle("Sleep-Wake pre trauma") + ylab("Mean Activity") + xlab("Hour of Day (0-23)") + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors) + ylim(c(.1,.25))

post_line_grouped <- ggplot(full_data[full_data$time == "recall5",], aes(x=Hr, y=mean, group = condition, colour =condition)) + stat_summary(aes(group=condition), fun.y=mean, geom="line") + ggtitle("Sleep-Wake post trauma") + ylab("Mean Activity") + xlab("Hour of Day (0-23)") + scale_fill_manual(values=group.colors) + scale_color_manual(values=group.colors) + ylim(c(.1,.25))
```

Plot graphs!
```{r}
pre_line_by_id
post_line_by_id

pre_line_grouped
post_line_grouped
```

# Analysis of high freezers vs low freezers

```{r}
high_freeze_ids <- c("fl05a5", "fl05a7", "fl5a6", "sert26a3")

freeze_high_low <- rep(NA, dim(full_data)[1])
for (i in 1:dim(full_data)[1]){
  if (full_data$id[i] %in% high_freeze_ids){
    freeze_high_low[i] <- "high"
  } else {
    freeze_high_low[i] <- "low"
  }
}

full_data <- cbind(full_data, freeze_high_low)
View(full_data)
```

Make graphs!

something is off with the code. I will work on it later. 
```{r}
pre_line_by_id_freeze <- ggplot(full_data[full_data$time == "baseline",], aes(x=Hr, y=mean, group = id, colour = freeze_high_low)) + geom_line() + ggtitle("Sleep-Wake pre trauma") + ylab("Mean Activity") + xlab("Hour of Day (0-23)") + ylim(c(0,.25))

post_line_by_id_freeze <- ggplot(full_data[full_data$time == "recall5",], aes(x=Hr, y=mean, group = id, colour =freeze_high_low)) + geom_line() + ggtitle("Sleep-Wake post trauma") + ylab("Mean Activity") + xlab("Hour of Day (0-23)")+ ylim(c(0,.25))

pre_line_grouped_freeze <- ggplot(full_data[full_data$time == "baseline",], aes(x=Hr, y=mean, group = id, colour =freeze_high_low)) + stat_summary(aes(group=condition), fun.y=mean, geom="line") + ggtitle("Sleep-Wake pre trauma") + ylab("Mean Activity") + xlab("Hour of Day (0-23)") + ylim(c(0,.25))

post_line_grouped_freeze <- ggplot(full_data[full_data$time == "recall5",], aes(x=Hr, y=mean, group = condition, colour =freeze_high_low)) + stat_summary(aes(group=condition), fun.y=mean, geom="line") + ggtitle("Sleep-Wake post trauma") + ylab("Mean Activity") + xlab("Hour of Day (0-23)") + ylim(c(0,.25))
```

Plot graphs!
```{r}
pre_line_by_id_freeze
post_line_by_id_freeze

pre_line_grouped_freeze
post_line_grouped_freeze
```



